---
title: "World Democracy Report"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    theme: united
runtime: shiny
---

<script>
$("body").on("shown.bs.tab", "a[data-toggle='tab']", function(e) {
   Shiny.setInputValue("active_tab", $(e.target).parent().index() + 1);
})
</script>

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(DT)
library(leaflet)
library(sp)
library(rnaturalearth)
library(plotly)
library(ggplot2)

democracy_number <- read.csv('../data/number-democracies-age-bmr.csv')
democracy_age <- read.csv('../data/age-of-democracy-bmr.csv')

liberal_democracy_number <- read.csv('../data/number-liberal-democracies-age.csv')
liberal_democracy_age <- read.csv('../data/age-of-liberal-democracy-row.csv')

electoral_democracy_number <- read.csv('../data/number-electoral-democracies-age.csv')
electoral_democracy_age <- read.csv('../data/age-of-electoral-democracy.csv')
```

# {.sidebar}

```{r}
selectInput(
  "dem_show", label = "Choose dataset:",
  choices = c('democracy','liberal-democracy','electoral-democracy'), selected = 'democracy'
)

renderUI({
  if (input$active_tab == 1)
    selectInput(
      "table_show",
      label = "Choose table:",
      choices = c('age', 'number'),
      selected = 'age'
    )
})

renderUI({
  if(input$active_tab==2 || input$active_tab==3)
    if(input$dem_show == 'democracy') {
      sliderInput('map_years','Years:',min=1800,max=2020,value=2020)
    }
    else {
      sliderInput('map_years','Years:',min=1789,max=2022,value=2022)
    }
})

renderUI({
  if (input$active_tab == 3 || input$active_tab == 4)
    selectInput(
      "region_select",
      label = "Region:",
      choices = c('Africa', 'Asia','Europe','North America','South America','Oceania','World'),
      selected = 'World'
    )
})

```

# Table {data-icon="fa-table"}

```{r}
# rename columns
colnames(democracy_number)[c(4,5,6,7,8,9)] <- c('non-democracies',
                                                '1-18 years',
                                                '19-30 years',
                                                '31-60 years',
                                                '61-90 years',
                                                '91 years +')
colnames(liberal_democracy_number)[c(4,5,6,7,8,9,10,11)] <- c('close autocracies',
                                                              'electoral autocracies',
                                                              'electoral democracies',
                                                              '1-18 years',
                                                              '19-30 years',
                                                              '31-60 years',
                                                              '61-90 years',
                                                              '91 years +')
colnames(electoral_democracy_number)[c(4,5,6,7,8,9,10)] <- c('91 years +',
                                                             '61-90 years',
                                                             '31-60 years',
                                                             '19-30 years',
                                                             '1-18 years',
                                                             'electoral autocracies',
                                                             'close autocracies')

colnames(democracy_age)[4] <- 'age'
colnames(liberal_democracy_age)[4] <- 'age'
colnames(electoral_democracy_age)[4] <- 'age'

renderDataTable({
  if(input$dem_show == 'democracy' & input$table_show == 'age') {
    datatable(democracy_age, rownames = F, options = list(pageLength = 15))
  }
  else if(input$dem_show == 'democracy' & input$table_show == 'number') {
    datatable(democracy_number, rownames = F, options = list(pageLength = 15))
  }
  else if(input$dem_show == 'liberal-democracy' & input$table_show == 'age') {
    datatable(liberal_democracy_age, rownames = F, options = list(pageLength = 15))
  }
  else if(input$dem_show == 'liberal-democracy' & input$table_show == 'number') {
    datatable(liberal_democracy_number, rownames = F, options = list(pageLength = 15))
  }
  else if(input$dem_show == 'electoral-democracy' & input$table_show == 'age') {
    datatable(electoral_democracy_age, rownames = F, options = list(pageLength = 15))
  }
  else if(input$dem_show == 'electoral-democracy' & input$table_show == 'number') {
    datatable(electoral_democracy_number, rownames = F, options = list(pageLength = 15))
  }
})
```

# Visualization {data-icon="fa-map"}

```{r}
renderLeaflet({
if(input$dem_show == 'democracy') {
  democracy_age %>% filter(Year==as.numeric(input$map_years)) -> map
  ne_countries <- ne_countries(scale = "medium", returnclass = "sf")
  map_sp <- merge(ne_countries, map, by.x = "iso_a3", by.y = "Code", all.x = TRUE)
  map_sp <- map_sp[, c("name","iso_a3", "geometry", "age")]
  map_sp$age[map_sp$age == "no data"] <- -2
  map_sp$age[map_sp$age == "non-democracy"] <- -1
  map_sp$age[is.na(map_sp$age)] <- -2
  map_sp$age <- as.numeric(map_sp$age)
  
  map_sp <- sf::st_transform(map_sp, "+proj=longlat +datum=WGS84")
  
  leaflet(map_sp) %>%
    addTiles() %>%
    addPolygons(
      fillColor = ~colorBin(c('#F2EFE9','red','#9ECAE1','#4292C6','#2171B5','#1E69AF','#084594'), domain = c(-2,-1,0,18,30,60,90,221), bins = c(-2,-1,0,18,30,60,90,221))(map_sp$age),
      weight = 2,
      opacity = 1,
      color = "white",
      dashArray = "3",
      fillOpacity = 0.7,
      highlight = highlightOptions(
        weight = 5,
        color = "#666",
        dashArray = "",
        fillOpacity = 0.7,
        bringToFront = TRUE
      ),
      label = ~ paste0(name, ": ", ifelse(
        age == -2, 'no data', ifelse(age == -1, 'non-democracy', age)
      ))
    ) %>%
    addLegend(
      position = "bottomright",
      title = "Age",
      colors = c('#F2EFE9','red','#9ECAE1','#4292C6','#2171B5','#1E69AF','#084594'),
      labels = c("No data", "Non-democracy", "0-18 years", "18-30 years", "30-60 years", "60-90 years", "90-221 years"),
      values = ~age,
      opacity = 1
    )
}
else if(input$dem_show == 'liberal-democracy') {
  liberal_democracy_age %>% filter(Year==as.numeric(input$map_years)) -> map
  ne_countries <- ne_countries(scale = "medium", returnclass = "sf")
  map_sp <- merge(ne_countries, map, by.x = "iso_a3", by.y = "Code", all.x = TRUE)
  map_sp <- map_sp[, c("name","iso_a3", "geometry", "age")]
  map_sp$age[map_sp$age == "no data"] <- -4
  map_sp$age[map_sp$age == "electoral democracy"] <- -1
  map_sp$age[map_sp$age == "electoral autocracy"] <- -2
  map_sp$age[map_sp$age == "closed autocracy"] <- -3
  map_sp$age[is.na(map_sp$age)] <- -4
  map_sp$age <- as.numeric(map_sp$age)
  
  map_sp <- sf::st_transform(map_sp, "+proj=longlat +datum=WGS84")
  
  leaflet(map_sp) %>%
    addTiles() %>%
    addPolygons(
      fillColor = ~colorBin(c('#F2EFE9','red','#FF5C5D','#EDF8B1','#9ECAE1','#4292C6','#2171B5','#1E69AF','#084594'), domain = c(-4,-3,-2,-1,0,18,30,60,90,174), bins = c(-4,-3,-2,-1,0,18,30,60,90,174))(map_sp$age),
      weight = 2,
      opacity = 1,
      color = "white",
      dashArray = "3",
      fillOpacity = 0.7,
      highlight = highlightOptions(
        weight = 5,
        color = "#666",
        dashArray = "",
        fillOpacity = 0.7,
        bringToFront = TRUE
      ),
      label = ~ paste0(name, ": ", ifelse(age==-4,'No data',ifelse(age==-3,'Closed autocracy',ifelse(age==-2,'Electoral autocracy',ifelse(age==-1,'Electoral democracy',age)))))
    ) %>%
    addLegend(
      position = "bottomright",
      title = "Age",
      colors = c('#F2EFE9','red','#FF5C5D','#EDF8B1','#9ECAE1','#4292C6','#2171B5','#1E69AF','#084594'),
      labels = c("No data","Closed autocracy", "Electoral autocracy","Electoral democracy", "0-18 years", "18-30 years", "30-60 years", "60-90 years", "90-174 years"),
      values = ~age,
      opacity = 1
    )
}
else if(input$dem_show == 'electoral-democracy') {
  electoral_democracy_age %>% filter(Year==as.numeric(input$map_years)) -> map
ne_countries <- ne_countries(scale = "medium", returnclass = "sf")
map_sp <- merge(ne_countries, map, by.x = "iso_a3", by.y = "Code", all.x = TRUE)
map_sp <- map_sp[, c("name","iso_a3", "geometry", "age")]
map_sp$age[map_sp$age == "no data"] <- -3
map_sp$age[map_sp$age == "electoral autocracy"] <- -1
map_sp$age[map_sp$age == "closed autocracy"] <- -2
map_sp$age[is.na(map_sp$age)] <- -3
map_sp$age <- as.numeric(map_sp$age)

map_sp <- sf::st_transform(map_sp, "+proj=longlat +datum=WGS84")

leaflet(map_sp) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~colorBin(c('#F2EFE9','red','#FF5C5D','#9ECAE1','#4292C6','#2171B5','#1E69AF','#084594'), domain = c(-3,-2,-1,0,18,30,60,90,174), bins = c(-3,-2,-1,0,18,30,60,90,174))(map_sp$age),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = ~ paste0(name, ": ", ifelse(age==-3,'No data',ifelse(age==-2,'Closed autocracy',ifelse(age==-1,'Electoral autocracy',age))))
  ) %>%
  addLegend(
    position = "bottomright",
    title = "Age",
    colors = c('#F2EFE9','red','#FF5C5D','#9ECAE1','#4292C6','#2171B5','#1E69AF','#084594'),
    labels = c("No data","Closed autocracy", "Electoral autocracy", "0-18 years", "18-30 years", "30-60 years", "60-90 years", "90-174 years"),
    values = ~age,
    opacity = 1
  )
}
})
```

# Chart {data-icon="fa-chart-line"}

```{r}
renderPlotly({
  if(input$dem_show == 'democracy') {
    democracy_number %>% filter(Entity==input$region_select, Year==as.numeric(input$map_years)) -> selected_data
    selected_data <- selected_data[4:9]
    df_gathered <- gather(selected_data, key = "variable", value = "value")
    
    ggplot(df_gathered, aes(x = variable, y = value)) +
      geom_bar(stat = "identity") +
      labs(x = NULL, y = "country numbers",title = paste('Number of democracies by age',input$region_select,input$map_years, sep = ', ')) +
      theme(plot.title = element_text(hjust = 0.5))
  }
  else if(input$dem_show == 'liberal-democracy') {
    liberal_democracy_number %>% filter(Entity==input$region_select, Year==as.numeric(input$map_years)) -> selected_data
    selected_data <- selected_data[4:11]
    df_gathered <- gather(selected_data, key = "variable", value = "value")
    
    ggplot(df_gathered, aes(x = variable, y = value)) +
      geom_bar(stat = "identity") +
      labs(x = NULL, y = "country numbers", title = paste('Number of liberal democracies by age',input$region_select,input$map_years, sep = ', '))+
      theme(plot.title = element_text(hjust = 0.5))
  }
  else if(input$dem_show == 'electoral-democracy') {
    electoral_democracy_number %>% filter(Entity==input$region_select, Year==as.numeric(input$map_years)) -> selected_data
    selected_data <- selected_data[4:10]
    selected_data <- selected_data[, rev(names(selected_data))]
    df_gathered <- gather(selected_data, key = "variable", value = "value")
    
    ggplot(df_gathered, aes(x = variable, y = value)) +
      geom_bar(stat = "identity") +
      labs(x = NULL, y = "country numbers", title = paste('Number of electoral democracies by age',input$region_select,input$map_years, sep = ', ')) +
      theme(plot.title = element_text(hjust = 0.5))
  }
})
```

# Summary {data-icon="fa-pie-chart"}

```{r}
renderPlotly({
  if(input$dem_show == 'democracy') {
    democracy_number %>% filter(Entity==input$region_select) -> selected_data
    selected_data <- selected_data[3:9]
    selected_data$democracies <- apply(selected_data[3:7],1,sum)
    
    ggplot() + 
      geom_line(aes(x = selected_data$Year, y = selected_data$democracies, color = "democracies")) +
      geom_line(aes(x = selected_data$Year, y = selected_data$`non-democracies`, color = "non democracies")) +
      labs(x="Years", y="Countries") +
      ggtitle(paste('Politycal systems',input$region_select, sep = ', ')) +
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_color_manual(name="Political System", 
                         values=c("democracies"="#9ECAE1", "non democracies"="#FF5C5D"))
  }
  else if(input$dem_show == 'liberal-democracy') {
    liberal_democracy_number %>% filter(Entity==input$region_select) -> selected_data
    selected_data <- selected_data[3:11]
    selected_data$`non-democracies` <- apply(selected_data[2:3],1,sum)
    selected_data$democracies <- apply(selected_data[4:9],1,sum)
    
    ggplot() + 
      geom_line(aes(x = selected_data$Year, y = selected_data$democracies, color = "democracies")) +
      geom_line(aes(x = selected_data$Year, y = selected_data$`non-democracies`, color = "non democracies")) +
      labs(x="Years", y="Countries") +
      ggtitle(paste('Politycal systems',input$region_select, sep = ', ')) +
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_color_manual(name="Political System", 
                         values=c("democracies"="#9ECAE1", "non democracies"="#FF5C5D"))
  }
  else if(input$dem_show == 'electoral-democracy') {
    electoral_democracy_number %>% filter(Entity==input$region_select) -> selected_data
    selected_data <- selected_data[3:10]
    selected_data <- selected_data[, rev(names(selected_data))]
    
    selected_data$`non-democracies` <- apply(selected_data[1:2],1,sum)
    selected_data$democracies <- apply(selected_data[3:7],1,sum)
    
    ggplot() + 
      geom_line(aes(x = selected_data$Year, y = selected_data$democracies, color = "democracies")) +
      geom_line(aes(x = selected_data$Year, y = selected_data$`non-democracies`, color = "non democracies")) +
      labs(x="Years", y="Countries") +
      ggtitle(paste('Politycal systems',input$region_select, sep = ', ')) +
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_color_manual(name="Political System", 
                         values=c("democracies"="#9ECAE1", "non democracies"="#FF5C5D"))
  }
})
```
# Sources

## column

### Creators

Data published by: Our World In Data, Bastian Herre  
Article: https://ourworldindata.org/democracies-age  
App creator: Paweł Ciszewski  
App code: https://github.com/kardahim/WorldDemocracyReport  


### Democracy Dataset

Based on the classification and assessment by Boix et al. (2013). **Democracies are
understood here as political systems in which political leaders are elected under broad
rights for men to vote in free and fair elections.**


## column

### Liberal Deomocracy Dataset

Based on the criteria of the classification by Lührmann et al. (2018) and the assessment by
V-Dem’s experts. **Liberal democracies are understood here as political systems that hold
meaningful, free and fair, and multi-party elections, and their citizens have further individual
rights, are equal before the law, and the legislative and the courts constrain the executive.**

### Electoral Democracy Dataset

Based on the criteria of the classification by Lührmann et al. (2018) and the assessment by
V-Dem’s experts. **Electoral democracies are understood here as political systems that hold
meaningful, free and fair, and multi-party elections.**